---
title: "fennica tutorial"
author: Leo Lahti and Niko Ilomaki
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{fennica Markdown Vignette}
%\usepackage[utf8]{inputenc}
-->



Fennica R package
===========

Tools for processing Finnish bibliographic data.



## Installation

Install development version with:

```{r install2, eval=FALSE}
library(devtools)
install_github("ropengov/fennica")
```

Test the installation by loading the library:

```{r test, message=FALSE, warning=FALSE, eval=FALSE}
library(fennica)
```


We also recommend setting the UTF-8 encoding:

```{r locale, eval=TRUE}
Sys.setlocale(locale="UTF-8") 
```


## Name - gender mappings

Gender information from the Finnish Population register. All first
names for living Finnish citizens that live in Finland and
abroad. Only names with frequency n>10 are included.  Source:
avoindata.fi service and [Vaestorekisterikeskus
(VRK)](https://www.avoindata.fi/data/fi/dataset/none). Version:
3/2016, licensed with CC-BY 4.0.

Get the name-gender mappings with:

```{r gender, eval=TRUE}
library(fennica)
tab <- get_gender_fi()
library(knitr)
kable(head(tab), digits = 2)
```


## Licensing and Citations

This work can be freely used, modified and distributed under the 
[Two-clause BSD license](http://en.wikipedia.org/wiki/BSD\_licenses).

```{r citation, message=FALSE, eval=TRUE}
citation("fennica")
```

## Session info

This vignette was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```






