---
title: "Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828"
author: "Mikko Tolonen, Jani Marjanen, Hege Roivainen, Leo Lahti"
date: "`r Sys.Date()`"
output: 
  beamer_presentation
---


```{r 201606krakow-init, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Default time span for the slides
min.year <- 1488
max.year <- 1828

library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)

knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(fig.path = "slides_201606_Krakow/", dev="CairoPNG")
knitr::opts_chunk$set(fig.path = "slides_201606_Krakow/")

# Set locale
Sys.setlocale(locale="UTF-8") 

# Nice theme
theme_set(theme_bw(26))
```


```{r 201606krakow-init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Read data
fen0 <- readRDS("fennica.Rds")
kun0 <- readRDS("kungliga.Rds")

fenraw <- readRDS("fennica.raw.Rds")
kunraw <- readRDS("kungliga.raw.Rds")

fen <- fen0 %>% filter(publication_year >= min.year & publication_year <= max.year)
kun <- kun0 %>% filter(publication_year >= min.year & publication_year <= max.year)

# Full time span
df.full = bind_rows(fen0, kun0)
df.full$catalog = factor(c(rep("Fennica", nrow(fen0)), rep("Kungliga", nrow(kun0))))

# Selected time spam
df0 = bind_rows(fen, kun)
df0$catalog = factor(c(rep("Fennica", nrow(fen)), rep("Kungliga", nrow(kun))))
```
---

## From particular data to global perspective?

Idea: to study early modern knowledge production in general.

How? By combining library catalogue data (Fennica, Kungliga, ESTC, CERL)

What is important: to clean up the data (90% of the work) and understand how particular data collections have been formed.

Fail: take worldcat or similar combination of datasets and analyse it without cleaning it first.

---

## Data and method

Data:

- Fennica (Finnish National Bibliography), n-number of documents

- Kungliga (Royal Swedish Library's metadata), n-number of documents

Method:

- Scalable code in library catalogue data ecosystem (github)

- xxx

---

## Argument and findings

- Finnish knowledge production 1640-1828 can only be understood in Swedish context.

- Academy of Turku as part of the Swedish University system is the centre in Finland.

Developments that start after "the rock bottom" of 1700-1721

1) Changes in University curriculum reflected in published theses 

2) Religious publishing in Finnish language 

3) Economy (and related subject fields) published in Finnish 

4) Enlightenment in Sweden reflected in octavo sized book publishing

---

## Peaks and drops – role of publishers and contingency

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=5}
library(devtools)
install_github("ropengov/gisfin")

# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishingovertime-1.png
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)

df2 <- df0 %>% group_by(publication_decade, catalog) %>% summarise(n = n())
p <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = catalog)) +
     geom_point(aes(shape = catalog), size = 3) +     
     ylab("Title count (n)") + xlab("Publication year") #+  
     # ggtitle("Overall publishing activity")
print(p)
```

---

## Comparing publishing activity 

- Turku
- Uppsala
- Lund 
- Stockholm 

```{r publishingactivitycomparisons, echo=FALSE, message=FALSE, cache=FALSE, fig.width=6, fig.height=5, fig.show="hold", out.width="150px"}
#https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishingactivitycomparisons-2.png
selected.places = c("Turku", "Uppsala", "Lund", "Stockholm")
df2 <- df0 %>% group_by(publication_decade, publication_place, catalog) %>%
               summarise(n = n()) %>%
	       filter(publication_place %in% selected.places) 
df2$catalog = factor(df2$catalog, levels = rev(c("Fennica", "Kungliga")))
df2$publication_place = droplevels(factor(df2$publication_place))
df3 = as.data.frame(spread(df2, publication_decade, n, fill = 0))
df2 = melt(df3)
colnames(df2) = c("publication_place", "catalog", "publication_decade", "n")
df2$publication_decade = as.numeric(as.character(df2$publication_decade))
df2$n = as.numeric(as.character(df2$n))
for (catal in unique(df2$catalog)) {
  p <- ggplot(subset(df2, catalog == catal), aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = publication_place)) +
     geom_point(aes(shape = publication_place), size = 3) +
     ylab("Title count (n)") + xlab("Publication year") +  
     ggtitle(catal)
  print(p)

}
```

---
 
## Unique publishers in top publication places over time 

```{r publishers2-finland, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers2-finland-1.png
df <- df0
df <- df %>%
        filter(catalog == "Fennica") %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

# Top publication places
ntop <- 10
library(bibliographica)
top <- names(top(df, "publication_place", ntop))

df <- df %>%	
        filter(publication_place %in% top) %>%	
        select(publication_decade, publication_place, publisher)
	
npub <- unique(df) %>% group_by(publication_decade, publication_place) %>% tally()
npub$publication_place <- factor(npub$publication_place, levels = top)
theme_set(theme_bw(20))

p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +        
       ggtitle("Unique publishers in top publication places (Finland/Fennica)")
print(p)
```

---

```{r publishers2-kungliga, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers2-kungliga-1.png
df <- df0
catalogue <- "Kungliga"
df <- df %>%
        filter(catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

# Selected publication places
top <- c("Stockholm", "Lund", "Uppsala", "Greifswald")
df <- df %>%	
        filter(publication_place %in% top) %>%	
        select(publication_decade, publication_place, publisher)
	
npub <- unique(df) %>% group_by(publication_decade, publication_place) %>% tally()
npub$publication_place <- factor(npub$publication_place, levels = top)
theme_set(theme_bw(20))

p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +        
       ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))
print(p)
```


---

## Title count per top publisher

```{r publishers3-finland-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=13, out.width="150px", fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers3-finland-1.png

place <- "Turku"
catalogue = "Fennica"
df = df0

df <- df %>%
        filter(catalog == catalogue & publication_place == place) 

df <- df %>%	
        select(publication_decade, publisher, paper.consumption.km2)

# Group small publishers
# Top publishers by title count
ntop <- 10
top <- names(top(df, "publisher", ntop))
df$publisher[!df$publisher %in% top] <- "Other"
df$publisher <- factor(df$publisher, c(top, "Other"))

# Title count per decade & publisher	
npub <- df %>% group_by(publication_decade, publisher) %>% tally()

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) + 
       xlab("Publication year") +
       ylab("Title count (n)") +        
       ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")", sep = ""))
print(p)


```

---

## Title count per top publisher

```{r publishers3-finland-kungliga, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=13, out.width="150px", fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers3-finland-1.png
catalogue = "Kungliga"
place <- "Turku"
df = df0

df <- df %>%
        filter(catalog == catalogue & publication_place == place) 

df <- df %>%	
        select(publication_decade, publisher, paper.consumption.km2)

# Group small publishers
# Top publishers by title count
ntop <- 10
top <- names(top(df, "publisher", ntop))
df$publisher[!df$publisher %in% top] <- "Other"
df$publisher <- factor(df$publisher, c(top, "Other"))

# Title count per decade & publisher	
npub <- df %>% group_by(publication_decade, publisher) %>% tally()

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) + 
       xlab("Publication year") +
       ylab("Title count (n)") +        
       ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")", sep = ""))
print(p)
```

---


## Publisher paper vs. title count 

Paper per title count ratio per publisher. [nähdään voidaanko sanoa onko joku selkeästi asetusten julkaisija ja toisella myös kirjoja (tai asetuskokoelmia).]

```{r publisherpapertitle, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5, out.width="150px", fig.show="hold"}

# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publisherpapertitle-1.png

for (catalogue in c("Fennica", "Kungliga")) {

  df <- df0
  df <- df %>%
        filter(catalog == catalogue) %>%
	filter(publication_year >= 1640 & publication_year <= 1828)

  # Selected ones
  df$publisher[df$publisher == ""] = "Other"
  top <- names(top(df, "publisher", 5))
  df$publisher[!df$publisher %in% c(top, "")] <- "Other"
  df$publisher <- factor(df$publisher, levels = c(top, "Other"))

  # Publishing per decade & publisher	
  npub <- df %>%
     	  group_by(publication_decade, publisher) %>%
	  summarize(paper = sum(paper.consumption.km2, na.rm = TRUE), n = n()) %>%
	  mutate(paper.per.title = paper/n)


theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = paper.per.title, group = publisher)) +
       geom_point(size = 5, aes(shape = publisher, color = publisher)) +
       geom_smooth(col = "black", aes(shape = publisher, fill = publisher, color = publisher)) + 
       xlab("Publication year") +
       ylab("Paper per title") +
       ylim(c(0, 1.1*max(npub$paper.per.title))) +
       ggtitle(paste("Paper per title ratio (", catalogue, ")", sep = ""))
print(p)
}
```

---


## Drop 1700

Publisher title count changes around 1700 drop in Turku, the most significant publishing place at the time:

```{r drop1700B, echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=7}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/drop1700B-1.png

catalogue <- "Fennica"
place <- "Turku"

df <- df0
df <- df %>%
        filter(catalog == catalogue & publication_place == place) %>%
        filter(publication_year >= 1650 & publication_year < 1750)

# Selected ones
top <- names(top(df, "publisher", 5))
df$publisher[!df$publisher %in% top] <- "Other"
df$publisher <- factor(df$publisher, levels = c(top, "Other"))

# Publishing per year & publisher	
npub <- df %>%
     	  group_by(publication_year, publisher) %>%
	  summarize(paper = sum(paper.consumption.km2, na.rm = TRUE), n = n())

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_year, y = n, group = publisher)) +
       geom_point(size = 5, aes(shape = publisher, color = publisher)) +
       geom_smooth(col = "black", aes(shape = publisher, fill = publisher, color = publisher)) + 
       xlab("Publication year") +
       ylab("Title count (n)") +
       ylim(c(0, 1.1*max(npub$n))) +
       ggtitle(paste("Title count (", place, "/", catalogue, ")", sep = ""))
print(p)
```

---

## Gävle in Riksdagar

An example how historical events, namely parliamentary assemblies in the Swedish case, have an impact on local knowledge production.

Gävle (gray) during riksdagar (Kungliga) TÄÄ EI TOIMI NYT OLLENKAAN KUN ON AIKA PIENI TOI NOUSU NÄKYVISSÄ TÄSSÄ. MENEEKÖ OIKEIN?

```{r riksdar2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/riksdar2-1.png

  # Use timeinterval year intervals
  myplace = "Stockholm" # Norrköping / Stockholm
  minyear = 1700
  maxyear = 1800
  catal = "Kungliga"

  df = subset(df0, publication_place == myplace & catalog == catal)
  df <- subset(df, publication_year >= minyear & publication_year < maxyear)

  timeinterval <- 1
  df$timeunit <- round(df$publication_year/timeinterval)*timeinterval 

  df$unity = rep(1, nrow(df))
  publications <- tapply(df$unity, list(df$timeunit), sum)

  publications[is.na(publications)] <- 0 # Set NAs to 0
  publications <- publications/timeinterval # Instead of decadal sum, use average annual output 
  dfm <- melt(publications) 
  names(dfm) <- c("Time", "Documents")
  dfm <- transform(dfm, date = as.numeric(as.character(Time)))
  ymin = min(dfm$Documents)
  ymax = max(dfm$Documents)

  rect_left <- c(min(na.omit(dfm$date)),
               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734
               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrköping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772
               1792-.5, 1792+.5, # Gävle 26 januari 1792 24 februari 1792	       
               max(na.omit(dfm$date)))
  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )
  rectangles$shade <- rep(c("White", "Highlight"), length = nrow(rectangles))

 riksplace = c("Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", "Gävle")
 cols = c("gray", "lightblue", "yellow", "white")
 rectangles$shade[rectangles$shade == "Highlight"] = riksplace
 rectangles$shade = factor(rectangles$shade)

  # Draw Figure
  theme_set(theme_bw(20))
  p <- ggplot()
  p <- p + geom_rect(data = rectangles, 
	   aes(xmin=xmin, xmax=xmax,
	       ymin=ymin, ymax=ymax,
	       fill=shade), alpha=0.8) + 
         scale_fill_manual(values = cols) # + guides(fill = "none") 
  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  #p <- p + scale_x_continuous(breaks = seq(minyear, maxyear, 20))
  p <- p + ggtitle("Publishing activity")
  p <- p + ylab("Documents / Year")
  p <- p + xlab("Year")
  p <- p + ggtitle(paste(myplace, " (", catal, ")", sep = ""))
  print(p)
```

---


# Language

## Book printing in Finland / Fennica

```{r language1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6, out.width="150px", fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/language1-1.png
catalogue = "Fennica"
df <- df0
langs <- c("Finnish", "Swedish", "Latin", "German", "Russian", "French", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
  
}

# ----------------------------------

df <- dfl %>% filter(catalog == catalogue)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# TITLE COUNT
p <- ggplot(df, aes(x = publication_decade, y = n, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)


# PAPER CONSUMPTION
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)
```

---

## Book printing in Finland / Kungliga

```{r language1-kungliga, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6, out.width="150px", fig.show="hold"}
catalogue = "Kungliga"
  df <- dfl %>% filter(catalog == catalogue)
  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))


# TITLE COUNT
p <- ggplot(df, aes(x = publication_decade, y = n, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)


# PAPER CONSUMPTION
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)
```


---

## Same with percentages

```{r language-perc, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6, out.width="150px", fig.show="hold"}
for (catalogue in c("Fennica", "Kungliga")) {

  if (catalogue == "Fennica") {
    df <- df.full # full time span
  } else if (catalogue == "Kungliga") {
    df <- df0 # selected time span
  }
  
  df <- df %>% filter(catalog == catalogue)

langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0

dfl <- NULL

for (lan in lang) {

  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
  
}

df <- dfl %>% group_by(publication_decade, language) %>%
     	     summarise(n = n())

# Calculate percentages
dff <- spread(df, language, n, fill = 0)
dff[, -1] <- 100 * t(apply(dff[, -1], 1, function (x) {x/sum(x)}))
dff <- gather(dff, publication_decade);
colnames(dff) <- c("publication_decade", "language", "f")
dff$language <- factor(dff$language, levels = langs)

theme_set(theme_bw(20))
p <- ggplot(dff, aes(x = publication_decade, y = f)) +
     geom_bar(position = "stack", stat = "identity", aes(fill = language)) + 
     xlab("Publication year") +
     ylab("Title count frequency (%)") +
     ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)


}
```

---
 

## Religious literature

Devotional literature (catechisms, hymns, prayer collections, etc.) is known to be the most important form of literature denoting the advancement of reading in Finland in nineteenth century. But when does it actually showup as a relevant category? 

```{r topics, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics-2.png
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
  
  # Selected catalogue with selected years
  df = filter(df0, catalog == catalogue & publication_year >= 1640 & publication_year <= 1828)

  # Selected topics
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
  df <- df %>% filter(catalog == catalogue & hit)

  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

  # PAPER CONSUMPTION
  theme_set(theme_bw(20))
  p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = "")) 
  print(p)
```

---


### BOOKINFINLAND DONE3

Temporary sermons played a crucial role in early modern knowledge production, but it has not been clearly analysed in what language they were in Finland (and Sweden). 


```{r topics2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics2-2.png

sel <- c("hautajaiset", "häät", "juhlamenot")
catalogue <- "Fennica"
  if (catalogue == "Fennica") {
    df <- df.full # full time span
  } else if (catalogue == "Kungliga") {
    df <- df0 # selected time span
  }
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  

# Selected catalogue with selected topics
df <- df %>% filter(catalog == catalogue & hit)

langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
}
dfl$language <- factor(dfl$language, levels = langs)

# -------------------------------------------

df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = n, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
#print(p)


# PAPER CONSUMPTION
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)
```

---

### BOOK IN FINLAND 4

Combination of devotional literature and temporary sermons gives us the following graphs.

```{r topics3, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics3-2.png

sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat","hautajaiset","häät","juhlamenot")
catalogue <- "Fennica"
  if (catalogue == "Fennica") {
    df <- df.full # full time span
  } else if (catalogue == "Kungliga") {
    df <- df0 # selected time span
  }
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  

# Selected catalogue with selected topics
df <- df %>% filter(catalog == catalogue & hit)

langs <- c("Finnish", "Swedish", "Latin", "German", "Russian", "French", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
}

# -------------------------------------------

df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = n, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
#print(p)


# PAPER CONSUMPTION
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)
```

---
 
# University teaching

### topics 1

Printing of theology related topics in different languages in Turku.

```{r topics-101, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel =  c("kirkkohistoria","raamatunhistoria","eksegetiikka","homiletiikka","dogmatiikka","teologia")
langs <- c("Finnish", "Swedish", "Latin", "German", "Russian", "French", "Other")
catalogue <- "Fennica"

df <- df0 %>% filter(catalog == catalogue)
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub) > 0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

### topics 2

Printing of history and related topics in different languages in Turku.

```{r topics-102, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("oppihistoria","antiikki","historia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

### topics 3

Printing of philosophy in different languages in Turku.

```{r topics-103, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

### topics 4

Printing of natural science in different languages in Turku.

```{r topics-104, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("luonnontieteet","matematiikka","fysiikka","kemia","tähtitiede","maantiede","geodesia","geofysiikka","meteorologia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

### topics 5

Printing of economy and related topics in different languages in Turku.


```{r topics-105, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel =   c("psykologia","yhteiskuntafilosofia","valtiofilosofia","talous","raha","moraali","velvollisuudet","kasvatus","maanviljely","maatalous","kalastus")

# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

# Enlightenment and reading

Rise of the octavo sized book as vehicle of Enlightenment (and proof of reading).

Paper consumption for different document formats over time. Each point represents a decade. Loess smoothing. 

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="150px"}
#https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/LIBER-13-2.png
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  p <- ggplot(df2, aes(y = paper.consumption.km2, x = publication_decade, shape = gatherings, linetype = gatherings))
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1, color = "black")
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + guides(linetype = guide_legend(keywidth = 5), shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", catal, ")", sep = ""))
  print(p)
}
```

---

## Turku
 
Effects of Enlightenment: decline of metaphysics in favour of social and economic theory and the great victory of reading is visible in the rise and rise of octavo book.

```{r 201606krakow-turkuvsother, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=5, fig.width=10}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/201606krakow-turkuvsother-2.png

df <- fen %>% 
      filter(!is.na(country)) %>%
      filter(!is.na(publication_place) & !is.na(paper.consumption.km2) & paper.consumption.km2 > -0.01)

# Add indicator field: Turku or Not Turku
myplace <- "Turku"

df$myplace <- rep(NA, nrow(df))
df$myplace[df$publication_place == myplace] <- myplace
df$myplace[!df$publication_place == myplace & df$country == "Finland"] <- "Finland"
df$myplace[!df$country == "Finland" & !df$publication_place == myplace] <- "Foreign"
df$myplace <- factor(df$myplace, levels = rev(c("Foreign", "Finland", "Turku")))

# Keep entries with known publication place and paper consumption
# Group by year/decade and Turku/Other
dfs <- df %>% group_by(publication_decade, myplace) %>%
  summarise(titles = n(), paper = sum(paper.consumption.km2, na.rm = TRUE))

p <- ggplot(dfs, aes(x = publication_decade, y = paper, fill = myplace, order = -as.numeric(myplace))) +
     geom_bar(position = "stack", stat = "identity", color = "black") +
     ggtitle("Paper consumption") +
     scale_fill_grey() +
     guides(fill = guide_legend(reverse = TRUE, title = "")) + 
     xlab("Publication decade") + ylab("Paper (km2)")
#print(p)


p <- ggplot(dfs, aes(x = publication_decade, y = titles, fill = myplace, order = -as.numeric(myplace))) +
     geom_bar(position = "stack", stat = "identity", color = "black") +
     ggtitle("Title count") +
     # scale_fill_manual(values = c("blue", "black", "darkgreen")) +
     scale_fill_grey() +
     guides(fill = guide_legend(reverse = TRUE, title = "")) +      
     xlab("Publication decade") + ylab("Title count (n)")
print(p)
```

---

## Thanks !

- Academy of Finland
- University of Helsinki
