---
title: "Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828"
author: "Mikko Tolonen, Jani Marjanen, Hege Roivainen, Leo Lahti"
date: "`r Sys.Date()`"
output: 
  beamer_presentation
---


```{r 201606krakow-init, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Default time span for the slides
min.year <- 1488
max.year <- 1828

library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)

knitr::opts_chunk$set(echo = FALSE)
#knitr::opts_chunk$set(fig.path = "slides_201606_Krakow/", dev="CairoPNG")
knitr::opts_chunk$set(fig.path = "slides_201606_Krakow/")

# Set locale
tmp = Sys.setlocale(locale="UTF-8") 

# Nice theme
theme_set(theme_bw(26))
```


```{r 201606krakow-init2, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
# Read data
fen0 <- readRDS("fennica.Rds")
kun0 <- readRDS("kungliga.Rds")

fenraw <- readRDS("fennica.raw.Rds")
kunraw <- readRDS("kungliga.raw.Rds")

fen <- fen0 %>% filter(publication_year >= min.year & publication_year <= max.year)
kun <- kun0 %>% filter(publication_year >= min.year & publication_year <= max.year)

# Full time span
df.full = bind_rows(fen0, kun0)
df.full$catalog = factor(c(rep("Fennica", nrow(fen0)), rep("Kungliga", nrow(kun0))))

# Selected time spam
df0 = bind_rows(fen, kun)
df0$catalog = factor(c(rep("Fennica", nrow(fen)), rep("Kungliga", nrow(kun))))
```

---

## From particular data to global perspective?

**Idea**: study early modern knowledge production, all of it (1470-1830)

**How?** By combining library catalogue data (Fennica, Kungliga, ESTC, CERL)

**Important**: to clean up the data (90% of the work) and to take into account how particular data collections (Finnish National Bibliography, English Short-Title Catalogue etc.) have been formed.

**How to fail**: take worldcat or similar combination of datasets and analyse it as a whole without cleaning it first.

---

[Intro 2/3]

### Data (1470-1828):

- Fennica (Finnish National Bibliography): 20328 documents

- Kungliga (National Library of Sweden): 73165 documents

### Method

- Scalable code in library catalogue data ecosystem (github)

- xxx TODO: WHAT DOES LEO WANT TO SAY HERE?

- Benefit no. 1: once the work has been done, the data is available for everyone to use and to enrich it. The ecosystem is constructed with even unforeseen changes in mind.

---

[Intro 3/3]

### Context

- Finnish knowledge production 1640-1828 can only be understood in the Swedish context.

- Academy of Turku as part of the Swedish University system is the centre of Finnish knowledge production.

### Findings

Enlightenment in Finland after "the rock bottom" of 1700-1721

1) Changes in University curriculum reflected in published theses 

2) Religious publishing in Finnish language 

3) Economy and social science published in Swedish language 

--> Deep roots of the diversity of "Finnishness" reflcted in knowledge production, 1640-1828.

---

[Big Picture 1/4]

### TODO Kartta!!! 

jossa näkyy Turku osana Ruotsin valtakuntaa 

ts. eri kaupungit ja niiden nimikemäärät kartalla. Kartat voisi tehdä, jos mahdollista, vuosille 1650, 1710, 1767 ja 1800 niin näkyy muutos. Tai sitten video jossa näkyy kehitys. Pohjois-Eurooppa karttapohjana.

Toteutuksesta riippuen voidaan miettiä, jos tämä kartta-slide korvaisi tuon alla tulevan "Comparing publishing activity in Swedish University town".

---

[Big Picture 2/4]

### Peaks and drops – role of publishers and contingency

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=5}
library(devtools)
install_github("ropengov/gisfin")

# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishingovertime-1.png
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)

df2 <- df0 %>% group_by(publication_decade, catalog) %>% summarise(n = n())
p <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = catalog)) +
     geom_point(aes(shape = catalog), size = 3) +     
     ylab("Title count (n)") + xlab("Publication year") #+  
     # ggtitle("Overall publishing activity")
print(p)
```

---

[Big Picture 3/4]

### Comparing publishing activity in Swedish University towns

- Turku
- Uppsala
- Lund 
- Stockholm 

```{r publishingactivitycomparisons, echo=FALSE, message=FALSE, cache=FALSE, fig.width=6, fig.height=5, fig.show="hold", out.width="150px"}
#https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishingactivitycomparisons-2.png
selected.places = c("Turku", "Uppsala", "Lund", "Stockholm")
df2 <- df0 %>% group_by(publication_decade, publication_place, catalog) %>%
               summarise(n = n()) %>%
	       filter(publication_place %in% selected.places) 
df2$catalog = factor(df2$catalog, levels = rev(c("Fennica", "Kungliga")))
df2$publication_place = droplevels(factor(df2$publication_place))
df3 = as.data.frame(spread(df2, publication_decade, n, fill = 0))
df2 = melt(df3)
colnames(df2) = c("publication_place", "catalog", "publication_decade", "n")
df2$publication_decade = as.numeric(as.character(df2$publication_decade))
df2$n = as.numeric(as.character(df2$n))
for (catal in unique(df2$catalog)) {
  p <- ggplot(subset(df2, catalog == catal), aes(x = publication_decade, y = n)) +
     geom_line(aes(linetype = publication_place)) +
     geom_point(aes(shape = publication_place), size = 3) +
     ylab("Title count (n)") + xlab("Publication year") +  
     ggtitle(catal)
  print(p)

}
```

---

[Big Picture 4/4]

### Rise of the octavo sized book as vehicle of Enlightenment

Paper consumption for different document formats over time. 

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8, fig.show="hold", out.width="150px"}
#https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/LIBER-13-2.png
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  p <- ggplot(df2, aes(y = paper.consumption.km2, x = publication_decade, shape = gatherings, linetype = gatherings))
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1, color = "black")
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + guides(linetype = guide_legend(keywidth = 5), shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", catal, ")", sep = ""))
  print(p)
}
```

---

### Publishers

---

[Publishers 1/6]

### TODO Turku in Sweden!!!

Janin ajatus: Kaaviot Fennicasta ja Kungligasta, jossa lasketaan eri kaupunkien nimikemäärien, paperinkulutuksen ja kustantajien suhteellinen osuus vuoteen 1828. Eli samanlainen kaavio kuin on kielistä ”Book printing in Finland” –osiossa. Laaditaan kaaviot Tukholmasta, Uppsalasta, Lundista, Greifswaldista ja Turusta. Voiko myös tehdä erillisen kaavion, jossa Tukholma on poistettu, josta näkee näiden kilpailevien kakkoskaupunkien tilanteen.

---

[Publishers 2/6]

### Development of number of publishers in Finland 

```{r publishers2-finland, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers2-finland-1.png
df <- df0
df <- df %>%
        filter(catalog == "Fennica") %>%
	filter(country == "Finland" | publication_place == "Vyborg") %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

# Top publication places
ntop <- 10
library(bibliographica)
top <- names(top(df, "publication_place", ntop))

df <- df %>%	
        filter(publication_place %in% top) %>%	
        select(publication_decade, publication_place, publisher)
	
npub <- unique(df) %>% group_by(publication_decade, publication_place) %>% tally()
npub$publication_place <- factor(npub$publication_place, levels = top)
theme_set(theme_bw(20))

p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +        
       ggtitle("Unique publishers in top publication places (Finland/Fennica)")
print(p)
```

---

[Publishers 3/6]

### Development of number of publishers in Swedish University towns

```{r publishers2-kungliga, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers2-kungliga-1.png
df <- df0
catalogue <- "Kungliga"
df <- df %>%
        filter(catalog == catalogue) %>%
        filter(publication_year >= 1640 & publication_year <= 1828)

# Selected publication places
top <- c("Stockholm", "Lund", "Uppsala", "Greifswald")
df <- df %>%	
        filter(publication_place %in% top) %>%	
        select(publication_decade, publication_place, publisher)
	
npub <- unique(df) %>% group_by(publication_decade, publication_place) %>% tally()
npub$publication_place <- factor(npub$publication_place, levels = top)
theme_set(theme_bw(20))

p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publication_place)) + 
       xlab("Publication year") +
       ylab("Unique publishers (n)") +        
       ggtitle(paste("Unique publishers in selected publication places (", catalogue, ")"))
print(p)
```


---

[Publishers 4/6]

### Title count of the output of publishers in Turku

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}

# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publishers3-finland-1.png

place <- "Turku"
catalogue = "Fennica"
df = df0

df <- df %>%
        filter(catalog == catalogue & publication_place == place)

df <- df %>%	
  filter(publication_year >= 1640 & publication_year <= 1828)%>%
        select(publication_decade, publisher, paper.consumption.km2)

# Group small publishers
# Top publishers by title count
ntop <- 10
top <- names(top(df, "publisher", ntop))
df$publisher[!df$publisher %in% top] <- "Other"
df$publisher <- factor(df$publisher, c(top, "Other"))

# Title count per decade & publisher	
npub <- df %>% group_by(publication_decade, publisher) %>% tally()

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) + 
       xlab("Publication year") +
       ylab("Title count (n)") +        
       ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")", sep = ""))
print(p)


```

---

[Publishers 5/6]

### Comparing paper used on average per publication by publishers


```{r publisherpapertitle, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, out.width="150px", fig.show="hold"}

# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/publisherpapertitle-1.png

for (catalogue in c("Fennica", "Kungliga")) {

  df <- df0
  df <- df %>%
        filter(catalog == catalogue) %>%
	filter(publication_year >= 1640 & publication_year <= 1828)

  # Selected ones
  df$publisher[df$publisher == ""] = "Other"
  top <- names(top(df, "publisher", 5))
  df$publisher[!df$publisher %in% c(top, "")] <- "Other"
  df$publisher <- factor(df$publisher, levels = c(top, "Other"))

  # Publishing per decade & publisher	
  npub <- df %>%
     	  group_by(publication_decade, publisher) %>%
	  summarize(paper = sum(paper.consumption.km2, na.rm = TRUE), n = n()) %>%
	  mutate(paper.per.title = paper/n)


theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = paper.per.title, group = publisher)) +
       geom_point(size = 5, aes(shape = publisher, color = publisher)) +
       geom_smooth(col = "black", aes(shape = publisher, fill = publisher, color = publisher)) + 
       xlab("Publication year") +
       ylab("Paper per title") +
       ylim(c(0, 1.1*max(npub$paper.per.title))) +
       ggtitle(paste("Paper per title ratio (", catalogue, ")", sep = ""))
print(p)
}
```

---

[Publishers 6/6]

### Rock Bottom of the Great Northern War (1700-1721)

Publisher title count changes around 1700 with a nearly detrimental drop in Turku, the most significant publishing place at the time.

```{r drop1700B, echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=7}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/drop1700B-1.png

catalogue <- "Fennica"
place <- "Turku"

df <- df0
df <- df %>%
        filter(catalog == catalogue & publication_place == place) %>%
        filter(publication_year >= 1650 & publication_year < 1750)

# Selected ones
top <- names(top(df, "publisher", 5))
df$publisher[!df$publisher %in% top] <- "Other"
df$publisher <- factor(df$publisher, levels = c(top, "Other"))

# Publishing per year & publisher	
npub <- df %>%
     	  group_by(publication_year, publisher) %>%
	  summarize(paper = sum(paper.consumption.km2, na.rm = TRUE), n = n())

theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_year, y = n, group = publisher)) +
       geom_point(size = 5, aes(shape = publisher, color = publisher)) +
       geom_smooth(col = "black", aes(shape = publisher, fill = publisher, color = publisher)) + 
       xlab("Publication year") +
       ylab("Title count (n)") +
       ylim(c(0, 1.1*max(npub$n))) +
       ggtitle(paste("Title count (", place, "/", catalogue, ")", sep = ""))
print(p)
```

---

[Swedish age of Freedom and abolishment of censorship 1/2]

### Gävle in Riksdagar

An example how historical events, namely parliamentary assemblies in the Swedish case, have an impact on local knowledge production.

Gävle (gray) during riksdagar (Kungliga) TODO: TÄÄ EI TOIMI NYT OLLENKAAN KUN ON AIKA PIENI TOI NOUSU NÄKYVISSÄ TÄSSÄ. MENEEKÖ OIKEIN? ERI NÄKÖINEN TUOLLA MANUSKRIPTIN PUOLELLA.

```{r riksdar2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/riksdar2-1.png

  # Use timeinterval year intervals
  myplace = "Stockholm" # Norrköping / Stockholm
  minyear = 1700
  maxyear = 1800
  catal = "Kungliga"

  df = subset(df0, publication_place == myplace & catalog == catal)
  df <- subset(df, publication_year >= minyear & publication_year < maxyear)

  timeinterval <- 1
  df$timeunit <- round(df$publication_year/timeinterval)*timeinterval 

  df$unity = rep(1, nrow(df))
  publications <- tapply(df$unity, list(df$timeunit), sum)

  publications[is.na(publications)] <- 0 # Set NAs to 0
  publications <- publications/timeinterval # Instead of decadal sum, use average annual output 
  dfm <- melt(publications) 
  names(dfm) <- c("Time", "Documents")
  dfm <- transform(dfm, date = as.numeric(as.character(Time)))
  ymin = min(dfm$Documents)
  ymax = max(dfm$Documents)

  rect_left <- c(min(na.omit(dfm$date)),
               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734
               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrköping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772
               1792-.5, 1792+.5, # Gävle 26 januari 1792 24 februari 1792	       
               max(na.omit(dfm$date)))
  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )
  rectangles$shade <- rep(c("White", "Highlight"), length = nrow(rectangles))

 riksplace = c("Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", "Gävle")
 cols = c("gray", "lightblue", "yellow", "white")
 rectangles$shade[rectangles$shade == "Highlight"] = riksplace
 rectangles$shade = factor(rectangles$shade)

  # Draw Figure
  theme_set(theme_bw(20))
  p <- ggplot()
  p <- p + geom_rect(data = rectangles, 
	   aes(xmin=xmin, xmax=xmax,
	       ymin=ymin, ymax=ymax,
	       fill=shade), alpha=0.8) + 
         scale_fill_manual(values = cols) # + guides(fill = "none") 
  p <- p + geom_line(data = dfm, aes(x = date, y = Documents), col = "black")
  p <- p + geom_point(data = dfm, aes(x = date, y = Documents), col = "black")
  #p <- p + scale_x_continuous(breaks = seq(minyear, maxyear, 20))
  p <- p + ggtitle("Publishing activity")
  p <- p + ylab("Documents / Year")
  p <- p + xlab("Year")
  p <- p + ggtitle(paste(myplace, " (", catal, ")", sep = ""))
  print(p)
```

---

[Swedish age of Freedom and abolishment of censorship 2/2]

### Number of publishers in Stockholm 1760-1770

TODO: Epäonnistuin tässä tuomaan koodin niin että knittaisi oikein. Täten joudun laittamaan kuvalinkin tähän. Harmittaa. Jotkut muut onnistui niin hyvin...

https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/polev1-3.png 

---

### Languages

TODO: "Other" kielikategorian selvittäminen seuraavissa kuvissa. Onko siellä joku tietty kieli joka näkyy hyvin selvästi mukana analyysissä? Nousee isoksi kategoriaksi. Nyt oon merkannut, että Kreikan kieli (Greek) mukaan moneen kohtaan. Jos kyseessä joku muu kuin Kreikka niin sitten se. Jos taas paljon kombinaatioita pieniä kieliä, niin antaa siinä tapauksessa olla.

---

[Languages 1/4]
#TODO: TÄSSÄ OLISI HYVÄ JOS v. 1800 julkaisut ei tippuisi nollaan!

### Development of use of different languages in Finnish National Bibliography

```{r language1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, out.width="160px", fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/language1-1.png
catalogue = "Fennica"
df <- df0
langs <- c("Finnish", "Swedish", "Latin", "German", "Russian", "French", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
  
}

# ----------------------------------

df <- dfl %>% filter(catalog == catalogue)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# TITLE COUNT
p <- ggplot(df, aes(x = publication_decade, y = n, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)


# PAPER CONSUMPTION
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)
```

---

[Languages 2/4]

### Development of use of different languages in Kungliga data

```{r language1-kungliga, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8, out.width="150px", fig.show="hold"}
catalogue = "Kungliga"
  df <- dfl %>% filter(catalog == catalogue) %>%
     	        group_by(publication_decade, language) %>%
                summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))


# TITLE COUNT
p <- ggplot(df, aes(x = publication_decade, y = n, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)


# PAPER CONSUMPTION
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)
```


---

[Languages 3/4]

### Different languages in Fennica and Kungliga (percentage)
TODO: GREEK mukaan (tai jos joku muu iso kieli niin se)!

```{r language-perc, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=12, out.width="150px", fig.show="hold"}
for (catalogue in c("Fennica", "Kungliga")) {

  if (catalogue == "Fennica") {
    df <- df.full # full time span
  } else if (catalogue == "Kungliga") {
    df <- df.full # full time span
  }
  
  df <- df %>% filter(catalog == catalogue)%>%
        filter(publication_year >= 1600 & publication_year <= 1900)

langs <- c("Finnish", "Swedish", "Latin", "German", "French", "Russian", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL

for (lan in lang) {

  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
  
}

df <- dfl %>% group_by(publication_decade, language) %>%
     	     summarise(n = n())

# Calculate percentages
dff <- spread(df, language, n, fill = 0)
dff[, -1] <- 100 * t(apply(dff[, -1], 1, function (x) {x/sum(x)}))
#dff <- gather(dff, publication_decade);
dff <- melt(dff, "publication_decade");
colnames(dff) <- c("publication_decade", "language", "f")
#dff$language <- factor(dff$language, levels = langs)

theme_set(theme_bw(20))
p <- ggplot(dff, aes(x = publication_decade, y = f)) +
     geom_bar(position = "stack", stat = "identity", aes(fill = language)) + 
     xlab("Publication year") +
     ylab("Title count frequency (%)") +
     ggtitle(paste("Languages (", catalogue, ")", sep = "")) +  scale_fill_brewer(palette="Paired")
print(p)
 
}
```

---

###TODO: KIELI% !!!!!!!!!!!!!

Sama prosenttianalyysi kielistä kuin edellisessä slidessä mutta paikkakuntakohtaisesti (sekä Fennican että Kungligan mukaan): Turku, Greifswald, Stockholm, Uppsala, Lund. Rajataan analyysi vuoteen 1828.

---

[Languages 4/4]

### Topic diversity in different languages in Finnish publications
TODO: TÄMÄ PITÄISI RAJATA LOPPUMAAN VUOTEEN 1828.

```{r topics232, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
df <- filter(df0, catalog == "Fennica") %>%
  filter(publication_year >= 1640 & publication_year <= 1828)%>% select(publication_year, language, subject_topic)
df <- unique(df)
df <- df %>% group_by(publication_year, language) %>% summarise(n = length(unique(unlist(strsplit(subject_topic, ";")))))

df$language[!df$language %in% langs] <- "Other"
df$language <- factor(df$language, levels = langs)

p <- ggplot(df, aes(x = publication_year, y = n, color = language)) +
       geom_line() + geom_point() +
       ggtitle("Number of subject topics")

print(p)
```

---

### Religious texts in Finland

---

[Religious texts in Finland 1/3]

### Devotional literature (catechisms, hymns, prayer collections, etc.) 

- Form of literature denoting the advancement of reading in Finland in the nineteenth century. It has been unclear when does it show up as a relevant category.

```{r topics12122, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6, fig.show="hold"}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics-2.png
sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat")
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
  
  # Selected catalogue with selected years
  df = filter(df0, catalog == catalogue & publication_year >= 1640 & publication_year <= 1828)

  # Selected topics
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
  df <- df %>% filter(catalog == catalogue & hit)

  lang <- paste("language.", langs, sep = "")
  otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
  df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
  dfl <- NULL
  for (lan in lang) {
    # Classify a document to the specifed language
    # If document is assigned with languages, each case is considered
    # so one doc may have multiple entries corresponding to different languages
    # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
    # Combine data frames for specified languages
    dflsub <- filter(df, df[[lan]])
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
  df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

  # PAPER CONSUMPTION
  theme_set(theme_bw(20))
  p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = "")) 
  print(p)
```

---

[Religious texts in Finland 2/3]
#TODO: TÄSTÄ ja seuraavasta TOI "IF FENNICA FULL TIME SPAN" KOMENTO POIS. En saanut koodattua sitä. TODO: GREEK mukaan!

### Temporary sermons in Latin and Swedish


- play an important role in early modern knowledge production. It has not been analysed in what language they were in Finland (and Sweden). 


```{r topics882, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics2-2.png

sel <- c("hautajaiset", "häät", "juhlamenot")
catalogue <- "Fennica"
  if (catalogue == "Fennica") {
    df <- df.full # full time span
  } else if (catalogue == "Kungliga") {
    df <- df0 # selected time span
  }
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  

# Selected catalogue with selected topics
df <- df %>% filter(catalog == catalogue & hit)

langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
}
dfl$language <- factor(dfl$language, levels = langs)

# -------------------------------------------

df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = n, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
#print(p)


# PAPER CONSUMPTION
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)
```

---

[Religious texts in Finland 3/3]
#TODO: GREEK mukaan!

### Combining devotional literature and temporary sermons


```{r topics3, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/topics3-2.png

sel <- c("virret","arkkiveisut","hartauskirjat","katekismukset","rukouspäivät","saarnat","aapiset","rukoukset","rukous","hengelliset laulut","hartauspuheet","virsikirjat","hautajaiset","häät","juhlamenot")
catalogue <- "Fennica"
  if (catalogue == "Fennica") {
    df <- df.full # full time span
  } else if (catalogue == "Kungliga") {
    df <- df0 # selected time span
  }
  df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  

# Selected catalogue with selected topics
df <- df %>% filter(catalog == catalogue & hit)

  # Selected catalogue with selected years
  df = filter(df0, catalog == catalogue & publication_year >= 1640 & publication_year <= 1828)

langs <- c("Finnish", "Swedish", "Latin", "German", "Other")
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  # Classify a document to the specifed language
  # If document is assigned with languages, each case is considered
  # so one doc may have multiple entries corresponding to different languages
  # mean(rowSums(df[, lang]) == 1) # 93% Fennica docs have just 1 language
  # Combine data frames for specified languages
  dflsub <- filter(df, df[[lan]])
  dflsub$language <- gsub("language.", "", lan)
  dfl <- bind_rows(dfl, dflsub)
}

# -------------------------------------------

df <- dfl %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = n, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Title count (n)") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
#print(p)


# PAPER CONSUMPTION
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Languages (", catalogue, ")", sep = ""))
print(p)
```

---

### University curriculum in Turku 1640-1828

---

[University curriculum 1/5]

TODO: OISKO NÄMÄ VIELÄ SELVEMPIÄ VÄRIKKÄINÄ BARPLOTTEINA? Leo, lisäksi säädätkö pois ton risuaitahässäkän näistä? En saanut sitä pois. TODO3: GREEK mukaan!

### Printing of theology in different languages in Turku

```{r topics-101, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel =  c("kirkkohistoria","raamatunhistoria","eksegetiikka","homiletiikka","dogmatiikka","teologia")
langs <- c("Finnish", "Swedish", "Latin", "German", "Russian", "French", "Other")
catalogue <- "Fennica"

df <- df0 %>% filter(catalog == catalogue)%>%
  filter(publication_year >= 1640 & publication_year <= 1800)
lang <- paste("language.", langs, sep = "")
otherlang <- setdiff(names(df)[grep("lang.", names(df))], lang)
df$language.Other <- rowSums(df[, otherlang] == TRUE, na.rm = T) > 0
dfl <- NULL
for (lan in lang) {
  dflsub <- filter(df, df[[lan]])
  if (nrow(dflsub) > 0) {
    dflsub$language <- gsub("language.", "", lan)
    dfl <- bind_rows(dfl, dflsub)
  }
}
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

[University curriculum 2/5]
TODO: GREEK mukaan!

### Printing of history in different languages in Turku


```{r topics-102, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("oppihistoria","antiikki","historia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

[University curriculum 3/5]

### Printing of philosophy in different languages in Turku
TODO: GREEK mukaan!

```{r topics-103, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

[University curriculum 4/5]
TODO: GREEK mukaan!

### Printing of natural science in different languages in Turku

```{r topics-104, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel = c("luonnontieteet","matematiikka","fysiikka","kemia","tähtitiede","maantiede","geodesia","geofysiikka","meteorologia")
# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

[University curriculum 5/5]
TODO: GREEK mukaan!

### Printing of economy and social science in different languages in Turku


```{r topics-105, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
sel =   c("psykologia","yhteiskuntafilosofia","valtiofilosofia","talous","raha","moraali","velvollisuudet","kasvatus","maanviljely","maatalous","kalastus")

# Selected catalogue with selected topics
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
	               paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_point(aes(col = language, shape = language), size = 5) +
       geom_line(aes(col = language, shape = language)) +       
       xlab("Publication year") +
       ylab("Paper consumption") +
       ggtitle(paste("Paper (", print(paste(sel, collapse = ";")), ")", sep = ""))
print(p)

```

---

### Effects of the Enlightenment in Finland

- decline of metaphysics, rise of social and economic theory 
- the Enlightenment literature read in Finland mainly printed elsewhere

TODO: KUVA PITÄIS RAJATA 1640-1828, EN ONNISTUNUT
```{r authors22, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5, out.width="150px", fig.show="hold"}
theme_set(theme_bw(20))
for (catalogue in c("Fennica", "Kungliga")) {
  df <- filter(df0, catalog == catalogue)
  top.authors <- names(top(df, field = "author", n = 10))
  dfs <- df %>% filter(author %in% top.authors) %>%
         group_by(author, publication_decade) %>%
         tally() %>%
         arrange(publication_decade)
  p <- ggplot(dfs, aes(x = publication_decade, y = n, fill = author)) +
       geom_bar(stat = "identity", position = "stack", color = "black") +
       xlab("Publication Decade") +
       ylab("Title Count") +
       scale_fill_grey() +
       guides(fill = guide_legend("Author")) +
       ggtitle(paste("Title count top authors\n", catalogue)) +
       xlim(range(df0$publication_year))
  print(p)
}
```

---

### Finland in the Swedish context

- Dramatic events regarding the diversity of "Finnishness" take place in the early nineteenth century --- but the analysis of Finnish knowledge production 1828-2016 is for another day.

```{r 201606krakow-turkuvsother, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=5, fig.width=10}
# https://github.com/rOpenGov/fennica/blob/master/inst/examples/figure_201606_Krakow/201606krakow-turkuvsother-2.png

df <- fen %>%
  filter(publication_year >= 1640 & publication_year <= 1800)%>% 
      filter(!is.na(country)) %>%
      filter(!is.na(publication_place) & !is.na(paper.consumption.km2) & paper.consumption.km2 > -0.01)

# Add indicator field: Turku or Not Turku
myplace <- "Turku"

df$myplace <- rep(NA, nrow(df))
df$myplace[df$publication_place == myplace] <- myplace
df$myplace[!df$publication_place == myplace & df$country == "Finland"] <- "Finland"
df$myplace[!df$country == "Finland" & !df$publication_place == myplace] <- "Foreign"
df$myplace <- factor(df$myplace, levels = rev(c("Foreign", "Finland", "Turku")))

# Keep entries with known publication place and paper consumption
# Group by year/decade and Turku/Other
dfs <- df %>% group_by(publication_decade, myplace) %>%
  summarise(titles = n(), paper = sum(paper.consumption.km2, na.rm = TRUE))

p <- ggplot(dfs, aes(x = publication_decade, y = paper, fill = myplace, order = -as.numeric(myplace))) +
     geom_bar(position = "stack", stat = "identity", color = "black") +
     ggtitle("Paper consumption") +
     scale_fill_grey() +
     guides(fill = guide_legend(reverse = TRUE, title = "")) + 
     xlab("Publication decade") + ylab("Paper (km2)")
#print(p)


p <- ggplot(dfs, aes(x = publication_decade, y = titles, fill = myplace, order = -as.numeric(myplace))) +
     geom_bar(position = "stack", stat = "identity", color = "black") +
     ggtitle("Title count") +
     # scale_fill_manual(values = c("blue", "black", "darkgreen")) +
     scale_fill_grey() +
     guides(fill = guide_legend(reverse = TRUE, title = "")) +      
     xlab("Publication decade") + ylab("Title count (n)")
print(p)
```

---

### Duplicate publications

TODO: General analysis (any simple picture) of duplicates between Kungliga and Fennica based on Hege's deduplication code.

---

### Thanks !

- Academy of Finland
- University of Helsinki
- National Library of Finland
- National Library of Sweden
- Digitalia / The Regional Council of South Savo
