---
title: "note_510c"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path=paste0("figure/rmd_", field, "_"))
```

* [Unique accepted titles](output.tables/title_accepted.csv): `r length(unique(df$note_510c))`
* [Unique discarded titles in original data](output.tables/title_discarded.csv) (excluding NA cases): `r as.vector(countLines(paste0(output.folder, "/note_510c_discarded.csv")) - 1)` 
* Original documents with non-NA titles: `r sum(!is.na(df$note_510c))` / `r nrow(df)` (`r round(100*mean(!is.na(df$note_510c)), 1)`%)
* Original documents with missing (NA) titles `r sum(is.na(df$note_510c))` / `r nrow(df)` documents (`r round(100*mean(is.na(df$note_510c)), 1)`%)


 Top-`r ntop` titles and their title counts.
 
```{r summarytitle, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
library(ggplot2)
theme_set(theme_bw(base_size = 12))
p <- top_plot(df, "note_510c", ntop = ntop, log10 = TRUE)+
  labs(y = "Entries (n)", title = "Top titles")
print(p)
``` 

Frequency of unique titles:
  
```{r uniquetitles, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}
tab <- table(df$note_510c)
d <- data.frame(title = names(tab), n = as.numeric(tab))

library(ggplot2)
theme_set(theme_bw(base_size = 12))
p <- ggplot(d, aes(x = n)) +
  geom_histogram(color="black", fill="white") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Title duplicates (n)", y = "Documents (n)")
print(p)
```
